An Embedding Framework for the Design and Analysis of Consistent Polyhedral Surrogates

### 2022-06-28 poly-embed-journal/poly-embed-journal.tex
If I keep saying it's getting closer, I'll be correct one of these days

\subsection{Hypographs of extended Bayes risks}\label{appsubsec:phase2}
\jessie{Remove subsection and merge with above?}

We now apply this polyhedral perspective to the Bayes risk of a loss function, extended to ``unnormalized distributions'', i.e., all of $\reals^\Y_+$.
Given a minimizable loss function $L : \R \to \reals^\Y_+$, define the $1$-homogeneous extension of its Bayes risk as
$\risk{L}_+:\reals^\Y\to\reals\cup\{\infty\}$, $x \mapsto \inf_{r\in\R} \inprod{x}{L(r)} - \delta(x \mid \reals^\Y_+)$.
In other words, letting $L(\R) := \{L(r) \mid r\in\R\} \subseteq \reals^\Y_+$, we have $\risk L_+ = g_{L(\R)}$.

%Using the polyhedral perspective above, we can characterize the hypograph of $\risk L_+ = g_{L(\R)}$ and extract a \emph{finite} set $\V$ of ``active'' loss vectors, so that $\risk L_+ = g_\V$.
%Let $\H_{L(\R)} = \{H_v^+ \mid v\in L(\R)\}$.
Observe that $L(\R)$ and $\H_{L(\R)}$ may be infinite sets.  
%We find the existence of a finite $\V \subseteq L(\R)$ such that $g_\V = g_{L(\R)}$, and proceed to work with such a $\V$.
\begin{claim}\label{claim:gV-equals-riskL}
	Suppose we are given a minimizable $L : \R \to \reals^\Y_+$ with polyhedral extended risk $\risk L_+$.
	Then $\hyp(g_{L(\R)}) = \cap(\H_\Y \cup \H_{L(\R)})$.
	%Moreover, there is a finite $\V \subseteq L(\R)$ such that $g_{L(\R)} = g_{\V}$. %%$\hyp(g_{L(\R)}) = \cap (\H_{\V} \cup \H_\Y) = \hyp(g_{\V})$ and . 
\end{claim}
\begin{proof}
	Observe that $x \in \reals^\Y_+ \iff (x, c) \in \cap \H_\Y$.
	Let $x \in \reals^\Y_+$.
	\begin{align*}
	(x,c) \in \hyp(g_{L(\R)})
	&\iff g_{L(\R)}(x) \geq c & \text{definition of hypograph}\\
	&\iff \risk L_+(x) \geq c & \text{$g_{L(\R)}=\risk L_+$}\\
    &\iff \inprod{v}{x} \geq c \,\, \forall v \in {L(\R)} & \text{definition of $\risk L_+$} \\
    % as the infimum % over $v \in {L(\R)}$ of}\\
    % & & \text{the inner product with $x$ and minimizable;} \\
	&\iff (x,c) \in H^+_{v} \,\, \forall v \in {L(\R)} & \text{definition of $H^+_v$}\\
	&\iff (x,c) \in \cap \H_{L(\R)} & 
	\end{align*}
	Combining the two equalities, we have $\hyp(g_{L(\R)}) = \cap (\H_\Y \cup \H_{L(\R)})$.
	% The existence of a finite $\V \subseteq {L(\R)}$ follows as $\risk L_+$ polyhedral implies $\hyp(\risk L_+)$ is a polyhedron, which has a finite representation by definition.

@@

  \jessiet{1. Show it's a facet. 2a. Show $H_y$ is a face. 2b. }
  To show that $\H_\Y \subseteq \H^*$, it suffices from Theorem~\ref{thm:polyhedron-uniquely-gen-facets}(ii) to show that each $F_y := H_y \cap \hyp(g_{L(\R)})$ is a facet.
  First, observe that $F_y$ is a face of $\hyp(g_{L(\R)})$ as $H_y$ supports $\hyp(g_{L(\R)})$ by construction of $g_{L(\R)}$ with the convex indicator. \jessie{too much hand-waving?}
  Moreover, $F_y = H_{(e_y, 0)} \cap \hyp(g_{L(\R)}) = \{(x,c) \mid x_y \geq 0\} \cap \hyp(g_{L(\R)})$ is a $|\Y|$-dimensional face as it contains $\{(x,c) \mid x_y \geq 0, c \leq 0\}$, which is $|\Y|$-dimensional.
%  To see $\H_\Y \subseteq \H^*$, if there was a $y \in \Y$ such that $H^+_y = \{(x,c) \mid x_y \geq 0\}$ was not in $\H^*$, it is either because it is redundant or $H^+_y \not \subseteq \hyp(g_{L(\R)})$.
%  In the first case, there must be some $\epsilon > 0$ such that $\{(x,c) \mid x_y \geq \epsilon\} \in \H^*$, which cannot happen as $g_{L(\R)}$ is concave and finite (namely, nonnegative) at $x = e_y$.
%  In the second case, we have a point $x \not \in \reals^\Y_+$ with $g_{L(\R)}(x) > -\infty$.
%  This contradicts the fact that $g_{L(\R)}$ is only finite on $\reals^\Y_+$ by construction with the convex indicator. \raft{TODO: Couldn't follow this last line, since $\delta$ can be $\infty$\jessie{does it help that it's minus the convex indicator?} 6/20/2022: still don't get it, sorry :-)... \jessie{Yeah I changed the sentence entirely.  better? \# burn}}. 
\end{proof}

@@

%\noindent
%The construction of $f_\V$ allows us to project $g_\V$ from $\reals^\Y_+$ to the $\simplex$.
%\begin{lemma}\label{cor:f-matches-g-on-simplex}
%  Consider a finite set $\V \subset \reals^\Y_+$.
%  For all $p \in \simplex$, we have $f_\V(p) = g_\V(p)$.
%\end{lemma}
%%Given the function $f_\V$, we consider $g_\V$ to be its extension and $L$ such that $\risk L_+ = g_\V$ and $\risk L = f_\V$.
Now, we define the function $\theta(v) = \{p \in \simplex \mid \inprod{v}{p} = f_\V(p)\}$ as the level sets of the loss vector $v \in \V$. %, and the set $\Theta_{\V^*} := \{\theta(v) \mid v \in \V^*\}$ to be the set of level sets uniquely defined by a given set of loss vectors.

@@

\begin{claim}\label{claim:theta-is-pi-cap-simplex}
  Consider $L, \V^*$ as in Assumption~\ref{assum:V-star-exists-infinite}.
  For all $v \in \V^*$, consider the face $F_v$ of $\hyp(g_{L(\R)})$.
  Then $\theta(v) = \pi(F_v) \cap \simplex$.
\end{claim}
\begin{proof}
  % Argue $\theta(v)$ is full-dim in $\simplex$ if and only if $\pi(F_v)$ is $|\Y|$-dim, if and only if $F_v$ is a facet. \raf{(2)}
  \btw{Moved previous proof outline to comment}

@@

\lemmaX*
\begin{proof}
	%\jessiet{Get to Assumption~\ref{assum:V-star-exists-infinite} quickly}
%As in this appendix, we consider $\V = L(\R)$ such that $g_\V = \risk L_+$.
%\jessie{$\risk L$ to $\risk L_+$.  Change Claim 6: if you have a minimizable loss and $\risk L$ is polyhedral, then $\risk L_+$ is polyhedral.}
Since $\risk L$ is polyhedral, so is $\risk L_+$ by Claim~\ref{claim:extended-poly-risk-poly}.
Therefore, we have satisfied the requirements of Proposition~\ref{prop:poly-risk-assumption-satisfied}, and can conclude there is a $\V$ such that $\risk L_+ = g_{\V}$ satisfying Assumption~\ref{assum:V-star-exists-infinite}.
Moreover, there must be a finite set $\R$ such that $\V = L(\R)$ by construction, which is representative as some element of $\R$ generates the Bayes risk for every $p \in \simplex$ as $\risk L = \min_{v \in L(\R)}\inprod{p}{v} - \delta\{p \mid \simplex\}$.
%Since $\risk L$ is polyhedral and concave there exists a finite set $\V$ such that $\risk L = f_\V$ by Claim~\ref{claim:f-min-affine}.
%Moreover as $\V$ is finite, $g_{\V} = \risk L_+$ (by $1$-homogeneity following from the form of $g_{\V}$) is also polyhedral and concave, and $f_\V = g_\V$ on $\simplex$.
%Therefore, there is a $\V^* \subseteq \V$ such that $L$ and $\V^*$ satisfy Assumption~\ref{assum:V-star-exists-infinite} by Proposition~\ref{prop:poly-risk-assumption-satisfied}.
%Moreover, any set $\R$ such that $\V^* = L(\R)$ is a finite representative set for $L$, and such a set must exist by the existence and construction of $\V^*$.
%%Given the form of $\risk L$ written as a pointwise max of finitely many affine functions (which are loss vectors) in $\V$, the loss $L$ has a finite representative set. 
%%Moreover, by Claim~\ref{claim:gV-and-fV}, $g_\V = f_\V$ on the simplex.

Now that we have the preliminaries of $L, \V^*$ satisfying Assumption~\ref{assum:V-star-exists-infinite}, consider the itemized statements.
Lemma~\ref{lem:lemX1-rep-iff-subset-vectors} is exactly statement \eqref{item:X-rep-V}.
This immediately implies statement~\eqref{item:X-min-V}.
Moreover, Lemma~\ref{lem:lemX-3-rep-iff-FDLS-subsets} is exactly statement~\eqref{item:X-rep-Theta}, and again statement~\eqref{item:X-min-Theta} immediately follows.
Statement~\eqref{item:X-rep-contain-min} is a corollary of the existence of a finite representative set, which follows since $\V \subseteq L(\R)$.
Statement~\eqref{item:X-full-dim} is exactly Lemma~\ref{lem:fdls-exactly-theta}.
Statement~\eqref{item:X-redundant} is exactly Lemma~\ref{lem:any-levelset-contained-in-minlevelset}.
Finally, Statement~\eqref{item:X-tight-embed} follows as a corollary of statement~\eqref{item:X-min-V} and Corollary~\ref{cor:tight-embed-min-rep}.
\end{proof}

>>>>

\subsection{Hypographs of extended Bayes risks}\label{appsubsec:phase2}
\noindent
We now apply this polyhedral perspective to the Bayes risk of a loss function, extended to ``unnormalized distributions'', i.e., all of $\reals^\Y_+$.
Given a minimizable loss function $L : \R \to \reals^\Y_+$, define the $1$-homogeneous extension of its Bayes risk as
$\risk{L}_+:\reals^\Y\to\reals\cup\{\infty\}$, $x \mapsto \inf_{r\in\R} \inprod{x}{L(r)} - \delta(x \mid \reals^\Y_+)$.
In other words, letting $L(\R) := \{L(r) \mid r\in\R\} \subseteq \reals^\Y_+$, we have $\risk L_+ = g_{L(\R)}$.
%
%Using the polyhedral perspective above, we can characterize the hypograph of $\risk L_+ = g_{L(\R)}$ and extract a \emph{finite} set $\V$ of ``active'' loss vectors, so that $\risk L_+ = g_\V$.
%Let $\H_{L(\R)} = \{H_v^+ \mid v\in L(\R)\}$.
Observe that $L(\R)$ and $\H_{L(\R)}$ may be infinite sets.  
%We find the existence of a finite $\V \subseteq L(\R)$ such that $g_\V = g_{L(\R)}$, and proceed to work with such a $\V$.
\begin{claim}\label{claim:gV-equals-riskL}
	Suppose we are given a minimizable $L : \R \to \reals^\Y_+$ with polyhedral extended risk $\risk L_+$.
	Then $\hyp(g_{L(\R)}) = \cap(\H_\Y \cup \H_{L(\R)})$.
	%Moreover, there is a finite $\V \subseteq L(\R)$ such that $g_{L(\R)} = g_{\V}$. %%$\hyp(g_{L(\R)}) = \cap (\H_{\V} \cup \H_\Y) = \hyp(g_{\V})$ and . 
\end{claim}
\begin{proof}
	Observe that $x \in \reals^\Y_+ \iff (x, c) \in \cap \H_\Y$.
	Let $x \in \reals^\Y_+$.
	\begin{align*}
	(x,c) \in \hyp(g_{L(\R)})
	&\iff g_{L(\R)}(x) \geq c & \text{definition of hypograph}\\
	&\iff \risk L_+(x) \geq c & \text{$g_{L(\R)}=\risk L_+$}\\
    &\iff \inprod{v}{x} \geq c \,\, \forall v \in {L(\R)} & \text{definition of $\risk L_+$} \\
    % as the infimum % over $v \in {L(\R)}$ of}\\
    % & & \text{the inner product with $x$ and minimizable;} \\
	&\iff (x,c) \in H^+_{v} \,\, \forall v \in {L(\R)} & \text{definition of $H^+_v$}\\
	&\iff (x,c) \in \cap \H_{L(\R)} & 
	\end{align*}
	Combining the two equalities, we have $\hyp(g_{L(\R)}) = \cap (\H_\Y \cup \H_{L(\R)})$.
	% The existence of a finite $\V \subseteq {L(\R)}$ follows as $\risk L_+$ polyhedral implies $\hyp(\risk L_+)$ is a polyhedron, which has a finite representation by definition.

@@

  To show that $\H_\Y \subseteq \H^*$, it suffices from Theorem~\ref{thm:polyhedron-uniquely-gen-facets}(ii) to show that each $F_y := H_y \cap \hyp(g_{L(\R)})$ is a facet.
  First, observe that $F_y$ is a face of $\hyp(g_{L(\R)})$ as $H_y$ supports $\hyp(g_{L(\R)})$ at the origin.\jessiet{Need to justify?}
  Moreover, $F_y = H_{(e_y, 0)} \cap \hyp(g_{L(\R)}) = \{(x,c) \mid x_y = 0\} \cap \hyp(g_{L(\R)})$ is a $|\Y|$-dimensional face as it contains $\{(x,c) \mid x_y = 0, c \leq 0\}$, which is $|\Y|$-dimensional.
%  To see $\H_\Y \subseteq \H^*$, if there was a $y \in \Y$ such that $H^+_y = \{(x,c) \mid x_y \geq 0\}$ was not in $\H^*$, it is either because it is redundant or $H^+_y \not \subseteq \hyp(g_{L(\R)})$.
%  In the first case, there must be some $\epsilon > 0$ such that $\{(x,c) \mid x_y \geq \epsilon\} \in \H^*$, which cannot happen as $g_{L(\R)}$ is concave and finite (namely, nonnegative) at $x = e_y$.
%  In the second case, we have a point $x \not \in \reals^\Y_+$ with $g_{L(\R)}(x) > -\infty$.
%  This contradicts the fact that $g_{L(\R)}$ is only finite on $\reals^\Y_+$ by construction with the convex indicator. \raft{TODO: Couldn't follow this last line, since $\delta$ can be $\infty$\jessie{does it help that it's minus the convex indicator?} 6/20/2022: still don't get it, sorry :-)... \jessie{Yeah I changed the sentence entirely.  better? \# burn}}. 
\end{proof}

@@

%\noindent
%The construction of $f_\V$ allows us to project $g_\V$ from $\reals^\Y_+$ to the $\simplex$.
%\begin{lemma}\label{cor:f-matches-g-on-simplex}
%  Consider a finite set $\V \subset \reals^\Y_+$.
%  For all $p \in \simplex$, we have $f_\V(p) = g_\V(p)$.
%\end{lemma}
%%Given the function $f_\V$, we consider $g_\V$ to be its extension and $L$ such that $\risk L_+ = g_\V$ and $\risk L = f_\V$.

@@

Now, we define the function $\theta(v) = \{p \in \simplex \mid \inprod{v}{p} = f_\V(p)\}$ as the level sets of the loss vector $v \in \V$. %, and the set $\Theta_{\V^*} := \{\theta(v) \mid v \in \V^*\}$ to be the set of level sets uniquely defined by a given set of loss vectors.
\begin{claim}\label{claim:theta-is-pi-cap-simplex}
  Consider $L, \V^*$ as in Assumption~\ref{assum:V-star-exists-infinite}.
  For all $v \in \V^*$, consider the face $F_v$ of $\hyp(g_{L(\R)})$.
  Then $\theta(v) = \pi(F_v) \cap \simplex$.
\end{claim}
\begin{proof}
  % Argue $\theta(v)$ is full-dim in $\simplex$ if and only if $\pi(F_v)$ is $|\Y|$-dim, if and only if $F_v$ is a facet. \raf{(2)}
  \btw{Moved previous proof outline to comment}

@@

\lemmaX*
\begin{proof}
	%\jessiet{Get to Assumption~\ref{assum:V-star-exists-infinite} quickly}
%As in this appendix, we consider $\V = L(\R)$ such that $g_\V = \risk L_+$.
%\jessie{$\risk L$ to $\risk L_+$.  Change Claim 6: if you have a minimizable loss and $\risk L$ is polyhedral, then $\risk L_+$ is polyhedral.}
Since $\risk L$ is polyhedral, so is $\risk L_+$ by Claim~\ref{claim:extended-poly-risk-poly}.
Therefore, we have satisfied the requirements of Proposition~\ref{prop:poly-risk-assumption-satisfied}, and can conclude there is a $\V$ such that $\risk L_+ = g_{\V}$ satisfying Assumption~\ref{assum:V-star-exists-infinite}.
Moreover, there must be a finite set $\R^*$ such that $\V = L(\R^*)$ by construction, which is representative as some element of $\R^*$ generates the Bayes risk for every $p \in \simplex$ as $\risk L = \min_{v \in L(\R^*)}\inprod{p}{v} - \delta\{p \mid \simplex\}$.
%Since $\risk L$ is polyhedral and concave there exists a finite set $\V$ such that $\risk L = f_\V$ by Claim~\ref{claim:f-min-affine}.
%Moreover as $\V$ is finite, $g_{\V} = \risk L_+$ (by $1$-homogeneity following from the form of $g_{\V}$) is also polyhedral and concave, and $f_\V = g_\V$ on $\simplex$.
%Therefore, there is a $\V^* \subseteq \V$ such that $L$ and $\V^*$ satisfy Assumption~\ref{assum:V-star-exists-infinite} by Proposition~\ref{prop:poly-risk-assumption-satisfied}.
%Moreover, any set $\R$ such that $\V^* = L(\R)$ is a finite representative set for $L$, and such a set must exist by the existence and construction of $\V^*$.
%%Given the form of $\risk L$ written as a pointwise max of finitely many affine functions (which are loss vectors) in $\V$, the loss $L$ has a finite representative set. 
%%Moreover, by Claim~\ref{claim:gV-and-fV}, $g_\V = f_\V$ on the simplex.

Now that we have the preliminaries of $L, \V$ satisfying Assumption~\ref{assum:V-star-exists-infinite}, consider the itemized statements.
Lemma~\ref{lem:lemX1-rep-iff-subset-vectors} is exactly statement \eqref{item:X-rep-V}.
This immediately implies statement~\eqref{item:X-min-V}.
Moreover, Lemma~\ref{lem:lemX-3-rep-iff-FDLS-subsets} is exactly statement~\eqref{item:X-rep-Theta}, and again statement~\eqref{item:X-min-Theta} immediately follows.
Statement~\eqref{item:X-rep-contain-min} is a corollary of the existence of a finite representative set, which follows since $\V \subseteq L(\R)$.
Statement~\eqref{item:X-full-dim} is exactly Lemma~\ref{lem:fdls-exactly-theta}.
Statement~\eqref{item:X-redundant} is exactly Lemma~\ref{lem:any-levelset-contained-in-minlevelset}.
Finally, Statement~\eqref{item:X-tight-embed} follows as a corollary of statement~\eqref{item:X-min-V} and Corollary~\ref{cor:tight-embed-min-rep}.
\end{proof}
